MJSLAAM: Designing a Sound, Configurable, Efficient, and Pushdown Monadic Static~Analyzer for JavaScript

### 2014-12-13 cs252r-fa14-writeup/paper.tex
stuff

@@

\title{MJSAI: Designing a Sound, Configurable, Efficient, and Pushdown %
              Monadic Static~Analyzer for JavaScript}
\author{{David Darais and Daniel King}\\
        School of Engineering and Applied Sciences\\
        Harvard University}

@@

We present the design and implementation of a static analyzer for \js{}. We
leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to
create an analysis which is both configurable and correct by construction. We
also recast pushdown static analysis in terms of \aam{}. This recasting permits
us to also easily create a pushdown \maam{} analysis for \js{}.

@@

\begin{itemize}
\item Design a configurable static analysis for javascript using the
  Monadic Abstracting Abstract Machines (\maam{}, \cite{maam}) design process.
\item Implement the analysis using the \maam{} framework as implemented in Haskell.
\item Analyze the performance of many different analysis techniques, including
  pushdown analysis, against the JSAI \cite{jsai} benchmark suite.
\end{itemize}

We completed the first goal, part of the second, and none of the third.
%
We underestimated the time necessary to implement the semantics of \js{} in the
\maam{} Haskell framework.
%
We were unable to complete a full implementation of the \js{} spec and
therefore unable to run our interpreter on the full JSAI~\cite{jsai} benchmark
suite.
%
However, we were successful in running our analysis on programs which reside in
the subset of \js{} that we were able to implement.
%
We also spent a significant amoutn of time understanding pushdown analysis and
recasting it in the \aam{} and \maam frameworks.

@@

We now describe an AAM-style approach to pushdown static analysis for a standard CEK machine.
%
Recall that a CEK machine is a state machine with three components:
%
\begin{enumerate}
\item The control expression $(c : Exp)$
\item The execution environment $(\rho : Var \rightarrow Val)$
\item The continuation $(k : List Frame)$
\end{enumerate}
%
In AAM-style, we first design an instrument concrete semantics.
%
This instrumented concrete semantics will yield a semantically equivalent
concrete interpreter, but with implementation choices which, after abstraction,
will yield an analysis with different properties.
%
To instrument the semantics for a pushdown analysis we:
\begin{enumerate}
\item Break the recursion between environments and values a. la. AAM \cite{aam}
\item Break the recursion between cons-cells in the list type for continuations.
%
This step is also in the discipline of AAM \cite{aam}, although this particular
choice is not made in their original paper.
%
\item Add a memo-table which caches the results of the analysis.
%
This memo-table caches results \emph{invariant to the continuation present during evaluation}.
\end{enumerate}

From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
\begin{enumerate}
\item The control expression $(c : Exp)$ as before
\item The environment $(\rho : Var \rightarrow Addr)$
\item The store $(\sigma : Addr \rightarrow Val)$
\item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
\item The continuation pointer $(\kappa-ptr : KAddr)$
\end{enumerate}

\begin{proposition}
A step relation
\begin{equation*}
\mapsto^{PD} \in CESKSK* \times CESKSK* 
\end{equation*}
is definable which agrees with the standard semantics
\begin{equation*}
\mapsto \in CEK \times CEK 
\end{equation*}
\end{proposition}

@@

@@

We, the authors, found the entire translation process completely straightforward
and mechanical, albeit time consuming due to the size of \lambdajs{} when
including the primitive operators.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% LstListings

\usepackage{listings}
\lstset{
  frame=none,
  xleftmargin=2pt,
  stepnumber=1,
  numbers=left,
  numbersep=5pt,
  numberstyle=\ttfamily\tiny\color[gray]{0.3},
  belowcaptionskip=\bigskipamount,
  captionpos=b,
  escapeinside={*'}{'*},
  language=haskell,
  tabsize=2,
  emphstyle={\bf},
  commentstyle=\it,
  stringstyle=\mdseries\rmfamily,
  showspaces=false,
  keywordstyle=\bfseries\rmfamily,
  columns=flexible,
  basicstyle=\small\sffamily,
  showstringspaces=false,
  morecomment=[l]\%,
  mathescape=true,
}

@@

\title{MJSLAAM: Designing a Sound, Configurable, Efficient, and Pushdown %
              Monadic Static~Analyzer for JavaScript}
\author{{David Darais and Daniel King}\\
        School of Engineering and Applied Sciences\\
        Harvard University}

@@

We present the design and implementation of MJSLAAM, a static analyzer for \js{}. We
leverage the Monadic Abstracting Abstract Machines (\maam{}) technique to
create an analysis which is both configurable and correct by construction. We
also recast pushdown static analysis in terms of \aam{}. This recasting permits
us to also easily create a pushdown \maam{} analysis for \js{}.
We call our framework MJSLAAM for Monadic JavaScript Language Abstract Abstract
Machine.

@@

\begin{itemize}
\item Design a configurable static analysis for javascript using the
  Monadic Abstracting Abstract Machines (\maam{}) \cite{maam} design process.
\item Implement the analysis using the \maam{} framework as implemented in Haskell.
\item Analyze the performance of many different analysis techniques, including
  pushdown analysis, against the JSAI \cite{jsai} benchmark suite.
\end{itemize}

We completed the first goal, part of the second, and none of the third.
%
We underestimated the time necessary to implement the semantics of \js{} in the
\maam{} Haskell framework.
%
We were unable to complete a full implementation of the \js{} spec and
therefore unable to run our interpreter on the full JSAI~\cite{jsai} benchmark
suite.
%
However, we were successful in running our analysis on small programs which
reside in the subset of \js{} that we were able to implement.
%
We also spent a significant amount of time understanding pushdown analysis and
recasting it in the \aam{} and \maam frameworks.

@@

We now describe an AAM-style approach to pushdown static analysis for a standard CEK machine.
%
Recall that a CEK machine is a state machine with three components:
%
\begin{enumerate}
\item The control expression $(c : Exp)$
\item The execution environment $(\rho : Var \rightarrow Val)$
\item The continuation $(k : List(Frame))$
\end{enumerate}
%
In AAM-style we first design an instrument concrete semantics.
%
This instrumented concrete semantics will yield a semantically equivalent
concrete interpreter, but with implementation choices which, after abstraction,
yield an analysis with pushdown properties.
%
To instrument the semantics for a pushdown analysis we:
\begin{enumerate}
\item Break the recursion between environments and values a. la. AAM \cite{aam}
\item Break the recursion between cons-cells in the list type for continuations.
%
This step is also in the discipline of AAM \cite{aam}, although this particular
choice is not made in their original paper.
%
\item Add a memo-table which caches the results of the analysis.
%
This memo-table caches results \emph{invariant to the continuation present during evaluation}.
\end{enumerate}

From steps 1. and 2. we end up with a C-E-S-KS-K* machine with the following components:
\begin{enumerate}
\item The control expression $(c : Exp)$ as before
\item The environment $(\rho : Var \rightarrow Addr)$
\item The store $(\sigma : Addr \rightarrow Val)$
\item The continuation store $(\kappa\sigma : KAddr \rightarrow \mathcal{P}(Frame + \{Null\}))$
\item The continuation pointer $(\kappa{}ptr : KAddr)$
\end{enumerate}

\begin{proposition}
A step relation
\begin{equation*}
\mapsto^{K} \in CESKSK^* \times CESKSK^* 
\end{equation*}
is definable which agrees with the standard semantics
\begin{equation*}
\mapsto \in CEK \times CEK 
\end{equation*}
\end{proposition}

@@

\begin{lstlisting}
data Frame =
    LetK [(Name, Set AValue)] Name [(Name, TExp)] TExp
  | AppL [TExp]
  | AppR (Set AValue) [(Set AValue)] [TExp]
  | ObjK [(String, (Set AValue))] Name [(Name, TExp)]
    -- Array Dereferencing
  | FieldRefL TExp
  | FieldRefR (Set AValue)
    -- Array Assignment
  | FieldSetA TExp         TExp
  | FieldSetN (Set AValue) TExp
  | FieldSetV (Set AValue) (Set AValue)
    -- Property Deletion
  | DeleteL TExp
  | DeleteR (Set AValue)
    -- Fig 2. Mutable References
  | RefSetL TExp
  | RefSetR (Set AValue)
  | RefK
  | DeRefK
    -- Fig 8. Control Operators
  | IfK TExp TExp
  | SeqK TExp
  | WhileL TExp TExp
  | WhileR TExp TExp
  | LabelK Label
  | BreakK Label
  | TryCatchK TExp Name
  | TryFinallyL TExp
  | TryFinallyR (Set AValue)
  | ThrowK
    -- Fig 9. Primitive Operations
  | PrimOpK Op [(Set AValue)] [TExp]
  deriving (Eq, Ord)
instance PartialOrder Frame where pcompare = discreteOrder
\end{lstlisting}
\caption{The Frame, i.e., Continuation Datatype}
\end{figure}

\begin{lstlisting}
pushFrame :: (Analysis $\varsigma$ m) => Frame -> m ()
pushFrame fr = do
  fp  <- getL konL
  fp' <- nextFramePtr
  modifyL kstoreL (mapInsertWith (\/) fp' (singleton (fr, fp))
  putL konL fp'

popFrame :: (Analysis $\varsigma$ m) => m Frame
popFrame = do
  fp <- getL konL
  k$\sigma$ <- getL kstoreL
  (fr, fp') <- mset (mjoin (liftMaybeSet (k$\sigma$ # fp)))
  putL konL fp'
  return fr
\end{lstlisting}
\caption{Pushing and Popping Stack Frames}
\end{figure}

@@

We found the entire translation process completely straightforward and
mechanical, albeit time consuming due to the size of \lambdajs{} and number of
primitive operations.
